---
title: "DS6306 - Analyzing US Craft Beers and Breweries - Case Study"
author: "Ahmad Salama & Almuhannad Qneis"
date: "2023-10-17"
output: html_document
---

## Introduction

Ladies and Gentlemen, CEO and CFO of Budweiser, my name is Ahmad, and I am joined by my college AL Muhannad.  

We are honored to present our data exploration and analysis of the Craft beers and breweries data you have provided. Our goal is to address your specific questions and report any noteworthy observations that can inform your decision-making. 

Now, let's dive into our data exploration and analysis. 

```{r load libraries}

# load required libraries
library(tidyverse)
library(dplyr)
library(GGally)
library(ggplot2)
library(ggthemes)
library(maps)
library(usdata)
library(class)
library(caret)
```

```{r prep}
# preparing and loading data related to US states and their geographical information
us_state_map_data <- map_data("state")
us_states <- data.frame(
  StateAbb = state.abb,
  StateName = tolower(state.name)
)
us_state_center_geo <- data.frame(region=tolower(state.name), CenterLong=state.center$x, CenterLatitude=state.center$y)
```
```{r set working directory}
#setwd("~/fearless")
```

```{r setting a seed for predictability}
# setting a seed for predictability
set.seed(4)
```


```{r loading breweries data}

#loading breweries data from csv
breweries_data_raw <- read.csv("data/Breweries.csv", header = TRUE, stringsAsFactors = FALSE)
# trim character values in the data and get rid of white space
breweries_data_cleaned <- breweries_data_raw %>% 
  sapply(function(x) if(is.character(x)) str_squish(x) else x)
breweries_data_cleaned <- as.data.frame(breweries_data_cleaned)
breweries_data_cleaned$Brew_ID <- as.integer(breweries_data_cleaned$Brew_ID)

# print summary of the breweries data set
summary(breweries_data_cleaned)
# print first 6 rows of the breweries data set
head(breweries_data_cleaned)
```

```{r load beers data}
# load beers data from csv:
beers_data_raw <-  as.data.frame(read.csv("data/Beers.csv", header = TRUE))
beers_data_cleaned <- beers_data_raw 
# change ABV from character to numeric values
beers_data_cleaned$ABV = as.numeric(beers_data_cleaned$ABV)
breweries_data_cleaned$State <- as.factor(str_squish(breweries_data_cleaned$State))

# print summary of the beers data set
summary(beers_data_cleaned)
# print first 6 rows of the beers data set
head(beers_data_cleaned)
```

```{r 1. Breweries in each state: }
# Q1.How many breweries are present in each state?
# 
# Before our analysis, we reviewed the brewery data set and verified that it had all the values and no missing values. This completeness ensures the reliability and accuracy of our insights, setting the stage for a robust analysis. 

# When it comes to craft breweries, our data suggest that it is Colorado that emerges as the champion with 47 breweries. California and Minnesota closely trail behind in the second and third positions, with 39 and 32 breweries, respectively. In the last place, we find Washington D.C., North Dakota, South Dakota, and West Virginia, each with only a single brewery. 

# verify no missing data.
breweries_missing_data_df <- breweries_data_cleaned %>% sapply(function(x) sum(is.na(x))) 



# creating a bar plot to visualize the count of breweries in different US states.
breweries_state_count <- breweries_data_cleaned %>% 
  group_by(State) %>%
  summarise(Count = n())

breweries_state_count %>%
  ggplot(mapping = aes(x = reorder(State, +Count))) +
  geom_bar(mapping = aes(y = Count), stat = "identity") + 
  ggtitle("Brewery Count by US State") +
  xlab("State") +
  ylab("Count") +
  geom_text(mapping = aes(x = reorder(State, +Count), y = Count, label = Count),
            stat = 'identity', position = position_dodge(.9), vjust = -0.5, size = 4, color = "red") +
  theme_economist()

# addition plot: create a heat map displaying the count of breweries in US states.
breweries_state_count <- merge(breweries_state_count,us_states, by.x = "State", by.y = "StateAbb", all.x = TRUE)
breweries_state_count <- left_join(us_state_map_data, breweries_state_count, by = c("region" = "StateName"))
breweries_state_count <- left_join(breweries_state_count, us_state_center_geo, by = c("region" = "region"))

breweries_state_count %>%
ggplot(mapping = aes(x = long, y = lat, group = group, fill = Count)) +
  geom_polygon() +
  geom_path(color = "white", size = 1.5) +  # Add state borders
  geom_text(mapping = aes(CenterLong, CenterLatitude, label = Count), color = "white") +
  coord_fixed(ratio=1.5) +
  scale_fill_gradient(low = "lightblue", high = "darkblue", 
                      name = "Brewery Count") +
  ggtitle("Brewery Count by US State") +
  xlab("Longitude") +
  ylab("Latitude") +
  theme_economist()
```


```{r 2.Merge beer data with the breweries data. Print the first 6 observations}

# merge breweries & beers data based on Brewery_id column.
beer_breweries <- left_join(beers_data_cleaned, breweries_data_cleaned, by = c("Brewery_id" = "Brew_ID"))
# change column names for clarity:
colnames(beer_breweries)[colnames(beer_breweries) == "Name.x"] <- "Beer_Name"
colnames(beer_breweries)[colnames(beer_breweries) == "Name.y"] <- "Brewery_Name"
head(beer_breweries)
```

```{r plot missing values function}
# function to create a bar plot to visualize the count of missing values in each column of a given data frame
plot_missing_values <- function(df) {
  # get the names of columns with missing values
  columns_with_missing_values <- names(which(sapply(df, function(x) sum(is.na(x)) > 0)))
  
  # calc missing values count for each column
  missing_values_count <- df %>%
    sapply(function(x) sum(is.na(x)))
  
  # create dataframe for columns with total number of mising values
  missing_values_df <- data.frame(
    column_name = columns_with_missing_values,
    missing_values_total = as.vector(sapply(columns_with_missing_values, function(x) {
                                              if(is.na(missing_values_count[x])) 0 else missing_values_count[x]
                                                 })))
  
  missing_values_df %>%
    ggplot(mapping = aes(x = column_name, y = missing_values_total)) +
    geom_bar(stat = "identity") + 
    ggtitle("Missing values Count by Column") +
    xlab("Missing Stat Column Name") +
    ylab("Count") +
    geom_text(mapping = aes(x = column_name, y = missing_values_total, 
                            label = missing_values_total),
              stat = 'identity', position = position_dodge(.9), vjust = -0.5, hjust = 2, size = 5) +
    geom_text(mapping = aes(x = column_name, y = missing_values_total, 
                            label = paste0(" (",scales::percent(missing_values_total/dim(beer_breweries)[1]), ")")),
              stat = 'identity', position = position_dodge(.9), vjust = -0.5, size = 5) +
    theme_economist()
}
```
```{r 3.Address the missing values in each column}
# the following chunk of code performs various data manipulation and imputation to handle missing values
# with plots to visualize missing values in each feature and visualize missing values by state.

# place holder for data with missing values
beer_breweries_with_missing_values <- beer_breweries

# calc mean for IBU and ABV before the handling of missing values
beer_breweries_summary_with_missing_values <- beer_breweries_with_missing_values %>% 
  summarize(IBU_Mean = mean(.data[["IBU"]], na.rm = TRUE),
            ABV_Mean  = mean(.data[["ABV"]], na.rm = TRUE),
            IBU_Median  = median(.data[["IBU"]], na.rm = TRUE), 
            ABV_Median  = median(.data[["ABV"]], na.rm = TRUE))

beer_breweries %>% sapply(function(x) sum(is.na(x))) 


missing_values_by_state <- beer_breweries %>% 
  group_by(State) %>%
  summarise(ibu_na = sum(is.na(IBU)), abv_na = sum(is.na(ABV)), count = n())

missing_values_by_state %>% ggplot(mapping = aes(x = reorder(State, +count))) +
  geom_bar(aes(y = count, fill = "Total Observations"), stat = "identity") +
  geom_bar(aes(y = ibu_na, fill = "IBU NA"), stat = "identity") +
  geom_bar(aes(y = abv_na, fill = "ABV NA"), stat = "identity") +
  scale_y_continuous(name = "Count", breaks = seq(0, 300, by = 20)) +
  ggtitle("Count of Missing Values in Total Observations by State") +
  xlab("State") +
  ylab("Count") +
  #scale_fill_identity(name = '', guide = 'legend',labels = c('IBV NA', 'IBU NA', 'Total Observations')) +
  scale_fill_manual(values = c("Total Observations" = "grey27", "IBU NA" = "forestgreen", "ABV NA" = "firebrick1")) +
  theme_economist()

plot_missing_values(beer_breweries)

# calc mean of ABV & IBU by city, state, and style:
abv_ibu_mean_grouped_by_style_city <- beer_breweries %>% 
  group_by(Style, City) %>%
  summarize(IBU = mean(.data[["IBU"]], na.rm = TRUE), ABV  = mean(.data[["ABV"]], na.rm = TRUE))

for(i in 1:nrow(beer_breweries)) {
  if(is.na(beer_breweries[i,"ABV"])) {
    sub_abv = abv_ibu_mean_grouped_by_style_city %>%
      filter(City == beer_breweries[i,"City"], Style == beer_breweries[i,"Style"])
    beer_breweries[i,"ABV"] <- sub_abv[1,"ABV"][1,1]
  }
  
  if(is.na(beer_breweries[i,"IBU"])) {
    sub_ibu = abv_ibu_mean_grouped_by_style_city %>% 
      filter(City == beer_breweries[i,"City"])
    beer_breweries[i,"IBU"] <- sub_ibu[1,"IBU"][1,1]
  }
}
plot_missing_values(beer_breweries)

beer_breweries %>% 
  summarize(IBU = mean(.data[["IBU"]], na.rm = TRUE), ABV  = mean(.data[["ABV"]], na.rm = TRUE))

abv_ibu_mean_grouped_by_city <- beer_breweries %>% 
  group_by(City) %>%
  select(City,ABV, IBU) %>%
  summarize_at(vars(ABV, IBU), mean, na.rm = TRUE)

for(i in 1:nrow(beer_breweries)) {
  if(is.na(beer_breweries[i,"ABV"])) {
    sub_abv = abv_ibu_mean_grouped_by_city %>%
      filter(City == beer_breweries[i,"City"])
    beer_breweries[i,"ABV"] <- sub_abv[1,"ABV"][1,1]
  }
  
  if(is.na(beer_breweries[i,"IBU"])) {
    sub_ibu = abv_ibu_mean_grouped_by_city %>% 
      filter(City == beer_breweries[i,"City"])
    beer_breweries[i,"IBU"] <- sub_ibu[1,"IBU"][1,1]
  }
}
plot_missing_values(beer_breweries)

beer_breweries %>% 
  summarize(IBU = mean(.data[["IBU"]], na.rm = TRUE), ABV  = mean(.data[["ABV"]], na.rm = TRUE))

abv_ibu_mean_grouped_by_state <- beer_breweries %>% 
  group_by(State) %>%
  select(State,ABV, IBU) %>%
  summarize_at(vars(ABV, IBU), mean, na.rm = TRUE)

for(i in 1:nrow(beer_breweries)) {
  if(is.na(beer_breweries[i,"ABV"])) {
    sub_abv = abv_ibu_mean_grouped_by_state %>%
      filter(State == beer_breweries[i,"State"])
    beer_breweries[i,"ABV"] <- sub_abv[1,"ABV"][1,1]
  }
  
  if(is.na(beer_breweries[i,"IBU"])) {
    sub_ibu = abv_ibu_mean_grouped_by_state %>% 
      filter(State == beer_breweries[i,"State"])
    beer_breweries[i,"IBU"] <- sub_ibu[1,"IBU"][1,1]
  }
}
beer_breweries %>% sapply(function(x) sum(is.na(x)))
plot_missing_values(beer_breweries)

#compare mean & median before and after fixing the values:
# calc mean for IBU and ABV before the handling of missing values
beer_breweries_summary_after_fixing_missing_values <- beer_breweries %>% 
  summarize(IBU_Mean = mean(.data[["IBU"]], na.rm = TRUE),
            ABV_Mean  = mean(.data[["ABV"]], na.rm = TRUE),
            IBU_Median  = median(.data[["IBU"]], na.rm = TRUE), 
            ABV_Median  = median(.data[["ABV"]], na.rm = TRUE))

# ABV mean before & after:
ggplot() +
  geom_boxplot(data = beer_breweries_with_missing_values,
               mapping = aes(x = 0, y = ABV)) +
  geom_point(data = beer_breweries_summary_with_missing_values,
             mapping = aes(x = 0, y = ABV_Mean), color = "red", shape = 18, size = 4) +
  geom_text(data = beer_breweries_summary_with_missing_values,
            mapping = aes(x = 0, y = ABV_Mean,
                          label = round(ABV_Mean, 3)), vjust = -0.75, color = "red") +
  geom_boxplot(data = beer_breweries,
               mapping = aes(x = 1, y = ABV)) +
  geom_point(data = beer_breweries_summary_after_fixing_missing_values,
             mapping = aes(x = 1, y = ABV_Mean), color = "red", shape = 18, size = 4) +
  geom_text(data = beer_breweries_summary_after_fixing_missing_values,
            mapping = aes(x = 1, y = ABV_Mean,
                          label = round(ABV_Mean, 3)), vjust = -0.75, color = "red") +
  ggtitle("ABV Distribution With & Without Missing Values Comparison") +
  xlab("") +
  ylab("ABV") +
  theme_economist()

# IBU mean before & after:
ggplot() +
  geom_boxplot(data = beer_breweries_with_missing_values,
               mapping = aes(x = 0, y = IBU)) +
  geom_point(data = beer_breweries_summary_with_missing_values,
             mapping = aes(x = 0, y = IBU_Mean), color = "red", shape = 18, size = 4) +
  geom_text(data = beer_breweries_summary_with_missing_values,
            mapping = aes(x = 0, y = IBU_Mean,
                          label = round(IBU_Mean, 3)), vjust = -0.75, color = "red") +
  geom_boxplot(data = beer_breweries,
               mapping = aes(x = 1, y = IBU)) +
  geom_point(data = beer_breweries_summary_after_fixing_missing_values,
             mapping = aes(x = 1, y = IBU_Mean), color = "red", shape = 18, size = 4) +
  geom_text(data = beer_breweries_summary_after_fixing_missing_values,
            mapping = aes(x = 1, y = IBU_Mean,
                          label = round(IBU_Mean, 3)), vjust = -0.75, color = "red") +
  ggtitle("IBU Distribution With & Without Missing Values Comparison") +
  xlab("") +
  ylab("IBU") +
  theme_economist()

```

```{r 4.Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare}
# calc median IBU & ABV by state
state_median_abv_ibu <- beer_breweries %>% 
  group_by(State) %>%
  select(State, ABV, IBU) %>%
  summarize_at(vars(ABV, IBU), median, na.rm = TRUE)

# create a bar plot for ABV median by state:
state_median_abv_ibu %>% 
  ggplot(mapping = aes(x = reorder(State, +ABV), y = ABV)) +
  geom_bar(stat = "identity") + 
  ggtitle("ABV Median by State") +
  xlab("State") +
  geom_text(mapping = aes(reorder(State, +ABV), y = ABV, 
                          label = paste0("             ", ABV), angle = 90),
            stat = 'identity', position = position_dodge(.9), size = 4, color = "red") +
  geom_hline(yintercept = median(state_median_abv_ibu$ABV, na.rm = TRUE), linetype = "dashed",
             color = "blue") +
  theme_economist()

# create a bar plot for IBU median by state:
state_median_abv_ibu %>% 
  ggplot(mapping = aes(x = reorder(State, +IBU), y = IBU)) +
  geom_bar(stat = "identity") + 
  ggtitle("IBU Median by State") +
  xlab("State") +
  geom_text(mapping = aes(reorder(State, +IBU), y = IBU, 
                          label = paste0("             ", round(IBU, 1)), angle = 90),
            stat = 'identity', position = position_dodge(.9), size = 4, color = "red") +
  geom_hline(yintercept = median(state_median_abv_ibu$IBU, na.rm = TRUE), linetype = "solid",
             color = "blue") +
  theme_economist()
```

```{r Maximum alcoholic (ABV) beer By State & State with max maximum bitter (IBU) beer}
# calculates and visualizes the maximum ABV & IBU by us state
# calc max ABV by state
max_abv_by_state <- beer_breweries %>% 
  group_by(State) %>%
  summarise(max_ABV = max(ABV, na.rm = TRUE)) %>%
  select(State, max_ABV)

# create a a bar plot for max ABV value in each state
max_abv_by_state %>% 
  ggplot(mapping = aes(x = reorder(State, +max_ABV))) +
  geom_bar(mapping = aes(y = max_ABV), stat = "identity") + 
  ggtitle("Max ABV by State") +
  xlab("State") +
  ylab("MAX ABV") +
  geom_text(mapping = aes(reorder(State, +max_ABV), y = max_ABV, 
                          label = paste0("             ", round(max_ABV, 4)), angle = 90),
            stat = 'identity', size = 4, color = "red") +
  theme_economist()

# calc max ibu by state
max_ibu_by_state <- beer_breweries %>% 
  group_by(State) %>%
  summarise(max_IBU = max(IBU, na.rm = TRUE)) %>%
  select(State, max_IBU)

# create a a bar plot for max IBU value in each state
max_ibu_by_state %>% 
  ggplot(mapping = aes(x = reorder(State, +max_IBU))) +
  geom_bar(mapping = aes(y = max_IBU), stat = "identity") + 
  ggtitle("Max IBU by State") +
  xlab("State") +
  ylab("MAX IBU") +
  geom_text(mapping = aes(reorder(State, +max_IBU), y = max_IBU, 
                          label = paste0("             ", round(max_IBU, 4)), angle = 90),
            stat = 'identity', size = 4, color = "red") +
  theme_economist()
```

```{r 6.summary statistics and distribution of the ABV variable.}

# calc summary of ABV values
summary(beer_breweries$ABV)

# create a box blot to show the distribution of ABV with average value
beer_breweries %>%
  ggplot(mapping = aes(y = ABV)) +
  geom_boxplot() +
  geom_point(aes(x = 0, y = mean(beer_breweries$ABV)), color = "red", shape = 18, size = 4) +
  geom_text(aes(x = 0, y = mean(beer_breweries$ABV), label = round(mean(beer_breweries$ABV), 3)), vjust = -0.75, color = "red") +
  ggtitle("ABV Distribution") +
  xlab("ABV") +
  ylab("Count") +
  theme_economist()

```

```{r 7.relationship between the bitterness of the beer and its alcoholic content}
# create visualizations for the relationship between ABV & IBU
beer_breweries %>%
  ggplot(mapping = aes(x = ABV, y = IBU)) +
  geom_point(na.rm = TRUE) +
  geom_smooth() +
  ggtitle("Alcohol Content (ABV) vs Bitterness (IBU)") +
  theme_economist()

# beer_breweries %>%
#   ggplot(mapping = aes(x = ABV, y = IBU)) +
#   geom_point(na.rm = TRUE) +
#   facet_grid(ABV + IBU ~ State) +
#   ggtitle("Alcohol Content (ABV) vs Bitterness (IBU)") +
#   theme_economist()

beer_breweries %>% 
  select(ABV, IBU) %>%
  ggpairs(columns = c("ABV", "IBU"))
```
```{r 8.Difference with respect to IBU and ABV between IPAs and ALE}
#Classify data by beer type
beer_breweries$Is_IPA = ifelse(str_detect(beer_breweries$Style, regex("IPA", ignore_case = TRUE)), TRUE, FALSE)
beer_breweries$Is_ALE = ifelse(!str_detect(beer_breweries$Style, regex("IPA", ignore_case = TRUE)) & (str_detect(beer_breweries$Beer_Name, regex("Ale", ignore_case = TRUE)) |
                                                                                                        str_detect(beer_breweries$Style, regex("Ale", ignore_case = TRUE))) , TRUE, FALSE)
### Create data set with beer type
beer_breweries_with_beer_type <- beer_breweries %>%
  filter( !(Is_ALE == FALSE & Is_IPA == FALSE), !(Is_ALE == TRUE & Is_IPA == TRUE)) %>%
  filter(!is.na(IBU)) %>%
  mutate(Beer_TYPE = as.factor(ifelse(Is_ALE == TRUE, "ALE", "IPA")))

# calc mean & median:
beer_type_abv_ibu_summary <- beer_breweries_with_beer_type %>% 
  group_by(Beer_TYPE) %>%
  summarize(ibu_mean = mean(IBU), ibu_median = median(IBU),
            abv_mean = mean(ABV), abv_median = median(ABV),
            abv_sd = sd(ABV), ibu_sd = sd(IBU))


# box plot to compare IBU distribution in ALE & IPA
beer_breweries_with_beer_type %>% 
  ggplot() +
  geom_boxplot(mapping = aes(x = Beer_TYPE, y = IBU)) +
  geom_point(data = beer_type_abv_ibu_summary[beer_type_abv_ibu_summary$Beer_TYPE == "IPA",],
             mapping = aes(x = "IPA", y = round(ibu_mean, 2), color = "Mean"), shape = 16, size = 4) +
  geom_text(data = beer_type_abv_ibu_summary[beer_type_abv_ibu_summary$Beer_TYPE == "IPA",],
            mapping = aes(x = "IPA", y = ibu_mean, label = round(ibu_mean, 2), color = "Mean"), vjust = -1.0) +
  geom_point(data = beer_type_abv_ibu_summary[beer_type_abv_ibu_summary$Beer_TYPE == "ALE",],
             mapping = aes(x = "ALE", y = round(ibu_mean, 2), color = "Mean"), shape = 16, size = 4) +
  geom_text(data = beer_type_abv_ibu_summary[beer_type_abv_ibu_summary$Beer_TYPE == "ALE",],
            mapping = aes(x = "ALE", y = ibu_mean, label = round(ibu_mean, 2), color = "Mean"), vjust = -1.0) +
  scale_y_continuous(breaks = seq(0, 150, by = 20)) +
  scale_fill_manual(values = c("Mean" = "firebrick1")) +
  ggtitle("IBU Distribution By Beer Type") +
  xlab("Beer Type") +
  theme_economist()

beer_breweries_with_beer_type %>% 
  ggplot() +
  geom_boxplot(mapping = aes(x = Beer_TYPE, y = ABV)) +
  geom_point(data = beer_type_abv_ibu_summary[beer_type_abv_ibu_summary$Beer_TYPE == "IPA",],
             mapping = aes(x = "IPA", y = round(abv_mean, 2), color = "Mean"), shape = 16, size = 4) +
  geom_text(data = beer_type_abv_ibu_summary[beer_type_abv_ibu_summary$Beer_TYPE == "IPA",],
            mapping = aes(x = "IPA", y = abv_mean, label = round(abv_mean, 2), color = "Mean"), vjust = -2) +
  geom_point(data = beer_type_abv_ibu_summary[beer_type_abv_ibu_summary$Beer_TYPE == "ALE",],
             mapping = aes(x = "ALE", y = round(abv_mean, 2), color = "Mean"), shape = 16, size = 4) +
  geom_text(data = beer_type_abv_ibu_summary[beer_type_abv_ibu_summary$Beer_TYPE == "ALE",],
            mapping = aes(x = "ALE", y = abv_mean, label = round(abv_mean, 2), color = "Mean"), vjust = -0.1) +
  scale_y_continuous(breaks = seq(0, 0.2, by = 0.01)) +
  scale_fill_manual(values = c("Mean" = "firebrick1")) +
  ggtitle("ABV Distribution By Beer Type") +
  xlab("Beer Type") +
  theme_economist()

beer_breweries_with_beer_type %>%
  ggplot(mapping = aes(x = ABV, y = IBU, color = Beer_TYPE)) +
  geom_point() +
  scale_y_continuous(breaks = seq(0, 160, by = 20)) +
  scale_x_continuous(breaks = seq(0, 0.8, by = 0.02)) +
  ggtitle("ABV & IBU Distribution By Beer Type") +
  xlab("IBU") +
  theme_economist()

#K-NN Modeling:
#scale ABV & IBU for k-NN model:
beer_breweries_with_beer_type$ABV_Scaled <- as.numeric((beer_breweries_with_beer_type$ABV - mean(beer_breweries_with_beer_type$ABV)) / sd(beer_breweries_with_beer_type$ABV))
beer_breweries_with_beer_type$IBU_Scaled <- as.numeric((beer_breweries_with_beer_type$IBU - mean(beer_breweries_with_beer_type$IBU)) / sd(beer_breweries_with_beer_type$IBU))
### Ploting Scaled ABV and IBU by beer type
beer_breweries_with_beer_type %>%
  ggplot(mapping = aes(x = ABV_Scaled, y = IBU_Scaled, color = Beer_TYPE)) +
  geom_point() +
  ggtitle("ABV Scaled & IBU Scaled Distribution By Beer Type") +
  xlab("ABV Scaled") +
  theme_economist()

###KNN after scaling
split_perc <- 0.70 
n <- dim(beer_breweries_with_beer_type)[1]
k_iterations <- 100
iterations_for_each_k <- 20

indices <- sample(1:n, size = round(split_perc * n))

beer_training_set <- beer_breweries_with_beer_type[indices,]
beer_testing_set <- beer_breweries_with_beer_type[-indices,]

# place holder for k & accuracy
accs <- data.frame(K = numeric(k_iterations), 
                   Accuracy = numeric(k_iterations),
                   Sensitiviy = numeric(k_iterations),
                   Specificity = numeric(k_iterations))

# iterate over different k values
for(i in 1:k_iterations) {
  accuracy <- 0
  sensitiviy <- 0
  specificity <- 0
  
  # loop through iterations with different and random training and testing sets.
  for(j in 1:iterations_for_each_k) {
    # randomly generate indices, by not setting the seed each iteration 
    # should be unique; hence, real cross validation
    indices <- sample(1:n, size = round(split_perc * n))
    beer_training_set <- beer_breweries_with_beer_type[indices,]
    beer_testing_set <- beer_breweries_with_beer_type[-indices,]
    # use knn, formula: Beer Type ~ ABV + IBU
    model <- knn(train = beer_training_set[, c("ABV_Scaled", "ABV_Scaled")], test = beer_testing_set[, c("ABV_Scaled", "ABV_Scaled")],
                 cl = beer_training_set$Beer_TYPE, k = i, prob = FALSE)
    # create the confusion matrix
    cm <- confusionMatrix(data = model, reference = beer_testing_set$Beer_TYPE)
    
    accuracy <- accuracy + cm$overall["Accuracy"]
    sensitiviy <- sensitiviy + cm$byClass["Sensitivity"]
    specificity <- specificity + cm$byClass["Specificity"]
  }
  
  # calculate average accuracy for the k valueL
  accs$K[i] <- i
  accs$Accuracy[i] <- round(accuracy / 20, 3)
  accs$Sensitiviy[i] <- round(sensitiviy / 20, 2)
  accs$Specificity[i] <- round(specificity / 20, 2)
}
#PLotting accuracy for different K iterations
accs %>% 
  ggplot(mapping = aes(x = accs$K)) +
  geom_line(y = accs$Accuracy, color = "firebrick") +
  geom_line(y = accs$Sensitiviy, color = "forestgreen") +
  geom_line(y = accs$Specificity, color = "blue")

# get the best accuracy row
best_accuracy <- filter(accs, accs$Accuracy == max(accs$Accuracy)) %>% top_n(1)

# plot the results and mark the best accuracy:
accs %>%
  ggplot(mapping = aes(x = K, y = Accuracy, type = "l", color = "Accuracy")) +
  geom_line() +
  geom_point(mapping = aes(x = best_accuracy$K, y = best_accuracy$Accuracy), color = "firebrick", show.legend = FALSE, size = 4, shape = 16) +
  scale_fill_manual(values = c("Accuracy" = "forestgreen")) +
  scale_x_continuous(breaks = seq(0, 100, by = 5)) +
  ggtitle("Beer Type KNN Model - Accuracy vs K Value") +
  theme_economist()
#KNN model with Optimal K value
# formula: Beer Type ~ ABV + IBU with K = 85, 
# indices <- sample(1:n, size = round(split_perc * n))
# beer_training_set <- beer_breweries_with_beer_type[indices,]
# beer_testing_set <- beer_breweries_with_beer_type[-indices,]
# model <- knn(train = beer_training_set[, c("ABV", "IBU")], test = beer_testing_set[, c("ABV", "IBU")],
#              cl = beer_training_set$Beer_TYPE, k = 85)
# # create a confusion matrix:
# cm <- confusionMatrix(data = model, reference = beer_testing_set$Beer_TYPE)
```

```{r 9. ABV & IBU Distribution By City in the State Of Colorado.}
#plot ABV and IBU by City in state of Colorado
beer_breweries_with_beer_type %>% 
  filter(State == "CO")%>% 
  group_by(City) %>% 
  ggplot(mapping = aes(x = ABV, y = IBU, color = City)) +
  geom_point() +
  scale_y_continuous(breaks = seq(0, 160, by = 20)) +
  ggtitle("ABV & IBU Distribution in CO") +
  xlab("ABV") +
  theme_economist()
```