---
title: "DS6306 - Analyzing US Craft Beers and Breweries - Case Study"
author: "Ahmad Salama & Almuhannad Qneis"
date: "2023-10-17"
output: html_document
---

## Introduction

Ladies and Gentlemen, CEO and CFO of Budweiser, we are honored to present to you our analysis of the data you've provided to us. Our objective is to uncover any exciting insights from the data you've provided, and we have focused our analysis on addressing specific questions while keeping the explanations straightforward and accessible.

Our analysis aims to explore the data and report any noteworthy observations that can inform your decision-making. This analysis provides an opportunity to understand your data better and discover valuable insights that may benefit Budweiser.

Now, let's delve into our analysis by addressing the questions you've outlined and any additional insights we've uncovered during our data exploration.

```{r setup, include=TRUE}

```

```{r load libraries}

# load libraries
library(tidyverse)
library(dplyr)
library(GGally)
library(ggplot2)
library(ggthemes)
library(maps)
library(usdata)
library(class)
library(caret)
```

```{r prep}
# preparing and loading data related to US states and their geographical information
us_state_map_data <- map_data("state")
us_states <- data.frame(
  StateAbb = state.abb,
  StateName = tolower(state.name)
)
us_state_center_geo <- data.frame(region=tolower(state.name), CenterLong=state.center$x, CenterLatitude=state.center$y)
```

```{r load breweries data}
# Set working directory
setwd("./data")
breweries_data_raw <- read.csv("Breweries.csv", header = TRUE, stringsAsFactors = FALSE)
# trim character values in the data
breweries_data_cleaned <- breweries_data_raw %>% 
  sapply(function(x) if(is.character(x)) str_squish(x) else x)
breweries_data_cleaned <- as.data.frame(breweries_data_cleaned)
breweries_data_cleaned$Brew_ID <- as.integer(breweries_data_cleaned$Brew_ID)

summary(breweries_data_cleaned)
head(breweries_data_cleaned)
```

```{r load beers data}
# Set working directory
setwd("./data")
# load beers data:
beers_data_raw <-  as.data.frame(read.csv("Beers.csv", header = TRUE))
beers_data_cleaned <- beers_data_raw 
# change ABV from character to numeric values
beers_data_cleaned$ABV = as.numeric(beers_data_cleaned$ABV)
breweries_data_cleaned$State <- as.factor(str_squish(breweries_data_cleaned$State))

summary(beers_data_cleaned)
head(beers_data_cleaned)
```

```{r 1. How many breweries are present in each state?}
# verify that there are no missing values in the breweries data set:
breweries_missing_data_df <- breweries_data_cleaned %>% sapply(function(x) sum(is.na(x))) #no missing data.

# creating a bar plot to visualize the count of breweries in different US states.
breweries_state_count <- breweries_data_cleaned %>% 
  group_by(State) %>%
  summarise(Count = n())

breweries_state_count %>%
  ggplot(mapping = aes(x = reorder(State, +Count))) +
  geom_bar(mapping = aes(y = Count), stat = "identity") + 
  ggtitle("Brewery Count by US State") +
  xlab("State") +
  ylab("Count") +
  geom_text(mapping = aes(x = reorder(State, +Count), y = Count, label = Count),
            stat = 'identity', position = position_dodge(.9), vjust = -0.5, size = 4, color = "red") +
  theme_economist()

# addition plot: create a heat map displaying the count of breweries in US states.
breweries_state_count <- merge(breweries_state_count,us_states, by.x = "State", by.y = "StateAbb", all.x = TRUE)
breweries_state_count <- left_join(us_state_map_data, breweries_state_count, by = c("region" = "StateName"))
breweries_state_count <- left_join(breweries_state_count, us_state_center_geo, by = c("region" = "region"))

breweries_state_count %>%
ggplot(mapping = aes(x = long, y = lat, group = group, fill = Count)) +
  geom_polygon() +
  geom_path(color = "white", size = 1.5) +  # Add state borders
  geom_text(mapping = aes(CenterLong, CenterLatitude, label = Count), color = "white") +
  coord_fixed(ratio=1.5) +
  scale_fill_gradient(low = "lightblue", high = "darkblue", 
                      name = "Brewery Count") +
  ggtitle("Brewery Count by US State") +
  xlab("Longitude") +
  ylab("Latitude") +
  theme_economist()
```


```{r 2.Merge beer data with the breweries data. Print the first 6 observations}

# merge breweries & beers data based on Brewery_id column.
beer_breweries <- left_join(beers_data_cleaned, breweries_data_cleaned, by = c("Brewery_id" = "Brew_ID"))
# change column names for clarity:
colnames(beer_breweries)[colnames(beer_breweries) == "Name.x"] <- "Beer_Name"
colnames(beer_breweries)[colnames(beer_breweries) == "Name.y"] <- "Brewery_Name"
head(beer_breweries)
```

```{r plot missing values function}
# function to create a bar plot to visualize the count of missing values in each column of a given data frame
plot_missing_values <- function(df) {
  # get the names of columns with missing values
  columns_with_missing_values <- names(which(sapply(df, function(x) sum(is.na(x)) > 0)))
  
  # calc missing values count for each column
  missing_values_count <- df %>%
    sapply(function(x) sum(is.na(x)))
  
  # create dataframe for columns with total number of mising values
  missing_values_df <- data.frame(
    column_name = columns_with_missing_values,
    missing_values_total = as.vector(sapply(columns_with_missing_values, function(x) {
                                              if(is.na(missing_values_count[x])) 0 else missing_values_count[x]
                                                 })))
  
  missing_values_df %>%
    ggplot(mapping = aes(x = column_name, y = missing_values_total)) +
    geom_bar(stat = "identity") + 
    ggtitle("Missing values Count by Column") +
    xlab("Missing Stat Column Name") +
    ylab("Count") +
    geom_text(mapping = aes(x = column_name, y = missing_values_total, 
                            label = missing_values_total),
              stat = 'identity', position = position_dodge(.9), vjust = -0.5, hjust = 2, size = 5) +
    geom_text(mapping = aes(x = column_name, y = missing_values_total, 
                            label = paste0(" (",scales::percent(missing_values_total/dim(beer_breweries)[1]), ")")),
              stat = 'identity', position = position_dodge(.9), vjust = -0.5, size = 5) +
    theme_economist()
}
```
```{r 3.Address the missing values in each column}
# the following chunk of code performs various data manipulation and imputation to handle missing values
# with plots to visualize missing values in each feature and visualize missing values by state.

# calc mean for IBU and ABV before the handling of missing values
beer_breweries %>% 
  summarize(IBU = mean(.data[["IBU"]], na.rm = TRUE), ABV  = mean(.data[["ABV"]], na.rm = TRUE))

beer_breweries %>% sapply(function(x) sum(is.na(x))) 

missing_values_count <- abv_ibu_mean_grouped_by_style_city %>%
  sapply(function(x) sum(is.na(x)))

missing_values_count <- abv_ibu_mean_grouped_by_style %>%
  sapply(function(x) sum(is.na(x)))

missing_values_by_state <- beer_breweries %>% 
  group_by(State) %>%
  summarise(ibu_na = sum(is.na(IBU)), abv_na = sum(is.na(ABV)), count = n())

missing_values_by_state %>% ggplot(mapping = aes(x = reorder(State, +count))) +
  geom_bar(aes(y = count, fill = "Total Observations"), stat = "identity") +
  geom_bar(aes(y = ibu_na, fill = "IBU NA"), stat = "identity") +
  geom_bar(aes(y = abv_na, fill = "ABV NA"), stat = "identity") +
  scale_y_continuous(name = "Count", breaks = seq(0, 300, by = 20)) +
  ggtitle("Count of Missing Values in Total Observations by State") +
  xlab("State") +
  ylab("Count") +
  #scale_fill_identity(name = '', guide = 'legend',labels = c('IBV NA', 'IBU NA', 'Total Observations')) +
  scale_fill_manual(values = c("Total Observations" = "grey27", "IBU NA" = "forestgreen", "ABV NA" = "firebrick1")) +
  theme_economist()

missing <- beer_breweries %>% 
  group_by(City) %>%
  summarise(sum_na = sum(is.na(IBU)), total = n())

plot_missing_values(beer_breweries)

# calc mean of ABV & IBU by city, state, and style:
abv_ibu_mean_grouped_by_style_city <- beer_breweries %>% 
  group_by(Style, City) %>%
  summarize(IBU = mean(.data[["IBU"]], na.rm = TRUE), ABV  = mean(.data[["ABV"]], na.rm = TRUE))

for(i in 1:nrow(beer_breweries)) {
  if(is.na(beer_breweries[i,"ABV"])) {
    sub_abv = abv_ibu_mean_grouped_by_style_city %>%
      filter(City == beer_breweries[i,"City"], Style == beer_breweries[i,"Style"])
    beer_breweries[i,"ABV"] <- sub_abv[1,"ABV"][1,1]
  }
  
  if(is.na(beer_breweries[i,"IBU"])) {
    sub_ibu = abv_ibu_mean_grouped_by_style_city %>% 
      filter(City == beer_breweries[i,"City"])
    beer_breweries[i,"IBU"] <- sub_ibu[1,"IBU"][1,1]
  }
}
plot_missing_values(beer_breweries)

beer_breweries %>% 
  summarize(IBU = mean(.data[["IBU"]], na.rm = TRUE), ABV  = mean(.data[["ABV"]], na.rm = TRUE))

abv_ibu_mean_grouped_by_city <- beer_breweries %>% 
  group_by(City) %>%
  select(City,ABV, IBU) %>%
  summarize_at(vars(ABV, IBU), mean, na.rm = TRUE)

for(i in 1:nrow(beer_breweries)) {
  if(is.na(beer_breweries[i,"ABV"])) {
    sub_abv = abv_ibu_mean_grouped_by_city %>%
      filter(City == beer_breweries[i,"City"])
    beer_breweries[i,"ABV"] <- sub_abv[1,"ABV"][1,1]
  }
  
  if(is.na(beer_breweries[i,"IBU"])) {
    sub_ibu = abv_ibu_mean_grouped_by_city %>% 
      filter(City == beer_breweries[i,"City"])
    beer_breweries[i,"IBU"] <- sub_ibu[1,"IBU"][1,1]
  }
}
plot_missing_values(beer_breweries)

beer_breweries %>% 
  summarize(IBU = mean(.data[["IBU"]], na.rm = TRUE), ABV  = mean(.data[["ABV"]], na.rm = TRUE))

abv_ibu_mean_grouped_by_state <- beer_breweries %>% 
  group_by(State) %>%
  select(State,ABV, IBU) %>%
  summarize_at(vars(ABV, IBU), mean, na.rm = TRUE)

for(i in 1:nrow(beer_breweries)) {
  if(is.na(beer_breweries[i,"ABV"])) {
    sub_abv = abv_ibu_mean_grouped_by_state %>%
      filter(State == beer_breweries[i,"State"])
    beer_breweries[i,"ABV"] <- sub_abv[1,"ABV"][1,1]
  }
  
  if(is.na(beer_breweries[i,"IBU"])) {
    sub_ibu = abv_ibu_mean_grouped_by_state %>% 
      filter(State == beer_breweries[i,"State"])
    beer_breweries[i,"IBU"] <- sub_ibu[1,"IBU"][1,1]
  }
}
beer_breweries %>% sapply(function(x) sum(is.na(x)))


plot_missing_values(beer_breweries)

# calc mean for IBU and ABV after the handling of missing values
beer_breweries %>% 
  summarize(IBU = mean(.data[["IBU"]], na.rm = TRUE), ABV  = mean(.data[["ABV"]], na.rm = TRUE))
```

```{r 4.Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare}
# calc median IBU & ABV by state
state_median_abv_ibu <- beer_breweries %>% 
  group_by(State) %>%
  select(State, ABV, IBU) %>%
  summarize_at(vars(ABV, IBU), median, na.rm = TRUE)

# create a bar plot for ABV median by state:
state_median_abv_ibu %>% 
  ggplot(mapping = aes(x = reorder(State, +ABV), y = ABV)) +
  geom_bar(stat = "identity") + 
  ggtitle("ABV Median by State") +
  xlab("State") +
  geom_text(mapping = aes(reorder(State, +ABV), y = ABV, 
                          label = paste0("             ", ABV), angle = 90),
            stat = 'identity', position = position_dodge(.9), size = 4, color = "red") +
  geom_hline(yintercept = median(state_median_abv_ibu$ABV, na.rm = TRUE), linetype = "dashed",
             color = "blue") +
  theme_economist()

# create a bar plot for IBU median by state:
state_median_abv_ibu %>% 
  ggplot(mapping = aes(x = reorder(State, +IBU), y = IBU)) +
  geom_bar(stat = "identity") + 
  ggtitle("IBU Median by State") +
  xlab("State") +
  geom_text(mapping = aes(reorder(State, +IBU), y = IBU, 
                          label = paste0("             ", round(IBU, 1)), angle = 90),
            stat = 'identity', position = position_dodge(.9), size = 4, color = "red") +
  geom_hline(yintercept = median(state_median_abv_ibu$IBU, na.rm = TRUE), linetype = "solid",
             color = "blue") +
  theme_economist()
```

```{r 5.Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?}
# calculates and visualizes the maximum ABV & IBU by us state
# calc max ABV by state
max_abv_by_state <- beer_breweries %>% 
  group_by(State) %>%
  summarise(max_ABV = max(ABV, na.rm = TRUE)) %>%
  select(State, max_ABV)

# create a a bar plot for max ABV value in each state
max_abv_by_state %>% 
  ggplot(mapping = aes(x = reorder(State, +max_ABV))) +
  geom_bar(mapping = aes(y = max_ABV), stat = "identity") + 
  ggtitle("Max ABV by State") +
  xlab("State") +
  ylab("MAX ABV") +
  geom_text(mapping = aes(reorder(State, +max_ABV), y = max_ABV, 
                          label = paste0("             ", round(max_ABV, 4)), angle = 90),
            stat = 'identity', size = 4, color = "red") +
  theme_economist()

# calc max ibu by state
max_ibu_by_state <- beer_breweries %>% 
  group_by(State) %>%
  summarise(max_IBU = max(IBU, na.rm = TRUE)) %>%
  select(State, max_IBU)

# create a a bar plot for max IBU value in each state
max_ibu_by_state %>% 
  ggplot(mapping = aes(x = reorder(State, +max_IBU))) +
  geom_bar(mapping = aes(y = max_IBU), stat = "identity") + 
  ggtitle("Max IBU by State") +
  xlab("State") +
  ylab("MAX IBU") +
  geom_text(mapping = aes(reorder(State, +max_IBU), y = max_IBU, 
                          label = paste0("             ", round(max_IBU, 4)), angle = 90),
            stat = 'identity', size = 4, color = "red") +
  theme_economist()
```

```{r 6.summary statistics and distribution of the ABV variable.}

# calc summary of ABV values
summary(beer_breweries$ABV)

# create a box blot to show the distribution of ABV with average value
beer_breweries %>%
  ggplot(mapping = aes(y = ABV)) +
  geom_boxplot() +
  geom_point(aes(x = 0, y = mean(beer_breweries$ABV)), color = "red", shape = 18, size = 4) +
  geom_text(aes(x = 0, y = mean(beer_breweries$ABV), label = round(mean(beer_breweries$ABV), 3)), vjust = -0.75, color = "red") +
  ggtitle("ABV Distribution") +
  xlab("ABV") +
  ylab("Count") +
  theme_economist()

```

```{r 7.relationship between the bitterness of the beer and its alcoholic content}
# create visualizations for the relationship between ABV & IBU
beer_breweries %>%
  select(ABV, IBU) %>%
  ggplot(mapping = aes(x = ABV, y = IBU)) +
  geom_point(na.rm = TRUE) +
  geom_smooth() +
  ggtitle("Alcohol Content (ABV) vs Bitterness (IBU)") +
  theme_economist()

beer_breweries %>% 
  select(ABV, IBU) %>%
  ggpairs(columns = c("ABV", "IBU"))
```