---
title: "DS6306 - Analyzing US Craft Beers and Breweries - Case Study"
author: "Ahmad Salama & Almuhannad Qneis"
date: "2023-10-17"
output: html_document
---

## Introduction

Ladies and Gentlemen, CEO and CFO of Budweiser, we are honored to present to you our analysis of the data you've provided to us. Our objective is to uncover any exciting insights from the data you've provided, and we have focused our analysis on addressing specific questions while keeping the explanations straightforward and accessible.

Our analysis aims to explore the data and report any noteworthy observations that can inform your decision-making. This analysis provides an opportunity to understand your data better and discover valuable insights that may benefit Budweiser.

Now, let's delve into our analysis by addressing the questions you've outlined and any additional insights we've uncovered during our data exploration.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load libraries
library(tidyverse)
library(dplyr)
library(GGally)
library(ggplot2)
library(ggthemes)
library(maps)
library(usdata)

# Set working directory
setwd("./data")

```
```{r load data}

```

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
